---
title: "Class 5: Data Visualization with ggplot"
author: "Andy (A12053233)"
format: pdf
---
sections 1-7
## Using ggplot
## To use ggplot2 we first need to install it on our computers. To do this, we will use the function `install.packages()`.

## Before I use any package functions, I have to load them up with a "library()" call, like so:

```{r}
library(ggplot2)
ggplot(cars)
```

```{r}
head(cars)
tail(cars)
```
## There is always the "base R" graphics system, i.e. the `plot()` function
## To use ggplot, I need to spell out at least 3 things: 
-data (the stuff I want to plot, always in the format of a data.frame)
-aesthetics (aes() values - how the data map to the plot)
-geometries/geoms (how I want things drawn)
```{r}
ggplot(cars) + 
  aes(x=speed, y=dist) + 
  geom_point() + 
  geom_smooth()
```

```{r}
ggplot(cars) + 
  aes(x=speed, y=dist) + 
  geom_point() + 
  geom_smooth(method="lm", se=FALSE) +
  labs(title="Speed and Stopping Distance of Cars",
       x="Speed (MPH)",
       y="Stopping Distance (ft)",
       subtitle="Your informative subtitle text",
       caption="Dataset: 'cars'") + 
  theme_bw()
```
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```
```{r}
nrow(genes)
colnames(genes)
ncol(genes)
sum(genes$State == "up")
```
There are `r nrow(genes)` in this dataset
There are `r sum(genes$State == "up")` "Up" regulated genes in this dataset
```{r}
table(genes$State)
round( table(genes$State)/nrow(genes) * 100, 2)
```
```{r}
ggplot(genes) + 
  aes(x=Condition1, y=Condition2) + 
  geom_point()
```
```{r}
p <- ggplot(genes) + 
  aes(x=Condition1, y=Condition2, col=State) + 
  geom_point()
p
p + scale_colour_manual( values=c("blue","gray","red")) + 
  labs(title="Gene Expression Changes Upon Drug Treatment",
       x="Control (no drug)",
       y="Drug Treatment")
```

```{r}

```

```{r}
#install.packages("gapminder")
library(gapminder)
#install.packages("dplyr")
library(dplyr)
gapminder_2007 <- gapminder %>% filter(year==2007)
ggplot(gapminder_2007) + 
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) + 
  geom_point(alpha=0.5)

```
```{r}
ggplot(gapminder_2007) + 
  aes(x=gdpPercap, y=lifeExp, color=pop) + 
  geom_point(alpha=0.8)
```
```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(x=gdpPercap, y=lifeExp, 
                 size=pop), alpha=0.5) +
  scale_size_area(max_size = 10)
  
```
```{r}
gapminder_1957 <- gapminder %>% filter(year==1957)

ggplot(gapminder_1957) + 
  aes(x = gdpPercap, y = lifeExp, color=continent, size = pop) +
  geom_point(alpha=0.7) + 
  scale_size_area(max_size = 10)
```
```{r}
gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)

ggplot(gapminder_1957) + 
  geom_point(aes(x=gdpPercap, y=lifeExp, color=continent,
                 size=pop), alpha=0.7) +
  scale_size_area(max_size = 15) + 
  facet_wrap(~year)
```

